---
title: "Data Mapping and Validation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data Mapping and Validation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Load Inputs

```{r setup}
library(fearbase)
library(readr)

data_long <- read_csv("data/data_long.csv", show_col_types = FALSE)
metadata <- read_csv("data/metadata.csv", show_col_types = FALSE)
```

## Mapping

`updateMapping()` loads the integrated condition-to-study mapping. By default it
also assigns `mapping` in the global environment for compatibility with older
interactive workflows. Use `assign_global = FALSE` in scripts when you want a
pure return value.

```{r update-mapping}
mapping <- updateMapping(assign_global = FALSE)
str(mapping)
```

Most public functions apply the relevant mapping internally, so normal users do
not need to call internal mapping helpers directly.

## Study Lists

`allStudies()` maps metadata and returns sorted study identifiers.

```{r all-studies}
studies <- allStudies(metadata)
head(studies)
length(studies)
```

## Strict Object Lookup

`checkData()` retrieves a named object from an environment and gives explicit
errors when the object cannot be found or when invalid arguments are supplied.
This is useful for OpenCPU-style workflows where object names arrive as strings.

```{r check-data}
example_data <- data.frame(x = 1:5, y = letters[1:5])

checkData("example_data")
checkData("example_data", inherits = TRUE)
```

## Diagnosing Removed Plot Rows

`traceRemovedRows()` helps identify rows that a ggplot layer is likely to drop
because required aesthetics are missing or coordinates fall outside the trained
panel range.

```{r trace-removed-rows}
plot_data <- data.frame(
  x = c(1, 2, NA, 100),
  y = c(1, 3, 4, 5)
)

plot <- ggplot2::ggplot(plot_data, ggplot2::aes(x, y)) +
  ggplot2::geom_point() +
  ggplot2::xlim(0, 10)

traceRemovedRows(plot, data = plot_data)
```

## OpenCPU CSV Loading

`createCsv()` reads an uploaded CSV file and returns a data frame. In local R
scripts, prefer `readr::read_csv()` or `utils::read.csv()` directly. In OpenCPU
workflows, upload the file to `createCsv()` first and pass the resulting session
object to downstream functions.

```{r create-csv}
uploaded_data <- createCsv("data/metadata.csv")
head(uploaded_data)
```

## JSON Summaries

`jsonSummary()` converts a named object to a JSON summary. The named object must
be visible to `checkData()`.

```{r json-summary}
jsonSummary("example_data")
```
