---
title: "FEARBASE Package Overview"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{FEARBASE Package Overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Purpose

The `fearbase` package provides analysis and visualization helpers for data
exported from the FEARBASE database. Most functions return `ggplot2` objects so
you can print, modify, combine, or save the result.

```{r setup}
library(fearbase)

data_long <- tibble::tribble(
  ~condition_id, ~study_id, ~paper_cond_id, ~paper_study_id, ~participant_id, ~measure, ~value, ~phase, ~stimulus, ~trial,
  "c1", "s1", "pc1", "ps1", "p1", "age", "21", "hab", "cs+", 1,
  "c1", "s1", "pc1", "ps1", "p1", "sex", "m", "hab", "cs+", 1,
  "c1", "s1", "pc1", "ps1", "p1", "scr", "0.3", "acq", "cs+", 1,
  "c1", "s1", "pc1", "ps1", "p2", "age", "24", "hab", "cs+", 1,
  "c1", "s1", "pc1", "ps1", "p2", "sex", "f", "hab", "cs+", 1,
  "c1", "s1", "pc1", "ps1", "p2", "us_exp", "40", "ext", "cs-", 1,
  "c2", "s2", "pc2", "ps2", "p3", "age", "29", "hab", "cs+", 1,
  "c2", "s2", "pc2", "ps2", "p3", "scr", "0.5", "acq", "cs+", 2,
  "c2", "s2", "pc2", "ps2", "p4", "age", "31", "ext", "cs-", 1
)

metadata <- tibble::tribble(
  ~condition_id, ~study_id, ~year, ~year_data, ~n_subjects, ~us_type, ~cs_type, ~instruction_contingency, ~reinf_acq, ~physio_scr_scoring_approach, ~physio_scr_baseline_window_start, ~physio_scr_baseline_window_end, ~physio_scr_peak_detection_window_min, ~physio_scr_peak_detection_window_max,
  "c1", "s1", 2020, 2019, 2, "electrotactile", "visual", "Fully instructed (whole exp)", 100, "baseline_correction", -1, 0, 1, 4,
  "c2", "s2", 2021, 2020, 2, "auditory", "visual", "Uninstructed (whole exp)", 75, "trough-to-peak", -2, 0, 2, 5
)

codebook <- tibble::tribble(
  ~attribute, ~abbreviation, ~name,
  "phase", "hab", "habituation",
  "phase", "acq", "acquisition",
  "phase", "ext", "extinction",
  "measure", "age", "age",
  "measure", "sex", "sex",
  "measure", "scr", "skin conductance response",
  "measure", "us_exp", "US expectancy"
)

study_design <- tibble::tribble(
  ~condition_id, ~study_id, ~name, ~cspTrials, ~csmTrials,
  "c1", "s1", "hab", 2, 2,
  "c1", "s1", "acq", 4, 4,
  "c2", "s2", "ext", 3, 3
)
```

## Main Data Inputs

The package works with four recurring FEARBASE tables:

- long-format participant data, usually passed as `dl`;
- study metadata, usually passed as `md`;
- the codebook, usually passed as `cb`;
- study-design data, usually passed as `sd`.

The package can also resolve common objects named `data_long`, `metadata`,
`codebook`, and `study_design` from the calling environment for interactive use.
Explicit arguments are preferred for reproducible scripts.

## Identifier Mapping

Many functions call internal mapping helpers before analysis. The mapping
normalizes legacy condition identifiers into current `condition_id` and
`study_id` columns.

```{r mapping}
mapping <- tibble::tibble(
  condition_id = c("c1", "c2"),
  study_id = c("s1", "s2")
)
head(mapping)

all_studies(metadata)
```

## Function Families

Participant summaries:

```{r participant-functions}
age(data_long, type = "histogram", grouping_variable = "study_id")
age(data_long, type = "ridge", grouping_variable = "condition_id")
age_descriptives(data_long, grouping_variable = "study_id")
sex(data_long)
sample_size_by_study(data_long)
```

Metadata summaries:

```{r metadata-functions}
data_collection_year(metadata)
instructions(metadata)
stimulus_modality(metadata, type = "us_type", level = "n_studies")
stimulus_modality(metadata, type = "cs_type", level = "n_subjects")
reinforcement_rates(metadata)
peak_detection_windows(metadata, grouping_variable = "study_id")
```

Trial and phase summaries:

```{r trial-functions}
trial_phase_counts(
  data_long,
  y_axis = "studies",
  grouping_variable = "condition_id",
  cb = codebook
)

trial_phase_counts(
  data_long,
  y_axis = "participants",
  grouping_variable = "study_id",
  cb = codebook
)
```

Heatmaps:

```{r heatmap-functions}
phases_heatmap(data_long, codebook)
measures_heatmap(data_long, metadata, codebook)
```

## Saving Results

Because plots are returned as objects, save them with `ggplot2::ggsave()`.

```{r save-plot}
plot <- sample_size_by_study(data_long)
# ggplot2::ggsave("sample-size-by-study.png", plot, width = 8, height = 5)
```
