---
title: "FEARBASE Package Overview"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{FEARBASE Package Overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Purpose

The `fearbase` package provides analysis and visualization helpers for data
exported from the FEARBASE database. Most functions return `ggplot2` objects so
you can print, modify, combine, or save the result.

```{r setup}
library(fearbase)
library(readr)

data_long <- read_csv("data/data_long.csv", show_col_types = FALSE)
metadata <- read_csv("data/metadata.csv", show_col_types = FALSE)
codebook <- read_csv("data/codebook.csv", show_col_types = FALSE)
study_design <- read_csv("data/study_design.csv", show_col_types = FALSE)
```

## Main Data Inputs

The package works with four recurring FEARBASE tables:

- long-format participant data, usually passed as `dl`;
- study metadata, usually passed as `md`;
- the codebook, usually passed as `cb`;
- study-design data, usually passed as `sd`.

The package can also resolve common objects named `data_long`, `metadata`,
`codebook`, and `study_design` from the calling environment for interactive use.
Explicit arguments are preferred for reproducible scripts.

## Identifier Mapping

Many functions call internal mapping helpers before analysis. The mapping
normalizes legacy condition identifiers into current `condition_id` and
`study_id` columns.

```{r mapping}
mapping <- updateMapping(assign_global = FALSE)
head(mapping)

allStudies(metadata)
```

## Function Families

Participant summaries:

```{r participant-functions}
age(data_long, type = "histogram", grouping_variable = "study_id")
age(data_long, type = "ridge", grouping_variable = "condition_id")
ageDescriptives(data_long, grouping_variable = "study_id")
sex(data_long)
sampleSizeByStudy(data_long)
```

Metadata summaries:

```{r metadata-functions}
dataCollectionYear(metadata)
instructions(metadata)
stimModality(metadata, type = "us_type", level = "n_studies")
stimModality(metadata, type = "cs_type", level = "n_subjects")
reinforcementRates(metadata)
peakDetectionWindows(metadata, grouping_variable = "study_id")
```

Trial and phase summaries:

```{r trial-functions}
trialsPhaseParticipant(
  data_long,
  y_axis = "studies",
  grouping_variable = "condition_id",
  cb = codebook
)

trialsPhaseParticipant(
  data_long,
  y_axis = "participants",
  grouping_variable = "study_id",
  cb = codebook
)
```

Heatmaps:

```{r heatmap-functions}
phasesHeatmap(data_long, codebook)
measuresHeatmap(data_long, metadata, codebook)
```

## Saving Results

Because plots are returned as objects, save them with `ggplot2::ggsave()`.

```{r save-plot}
plot <- sampleSizeByStudy(data_long)
ggplot2::ggsave("sample-size-by-study.png", plot, width = 8, height = 5)
```
