---
title: "Metadata Visualizations"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Metadata Visualizations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Load Metadata

```{r setup}
library(fearbase)
library(readr)

metadata <- read_csv("data/metadata.csv", show_col_types = FALSE)
```

## Data Collection Year

`dataCollectionYear()` counts studies or metadata records per valid year after
mapping metadata identifiers.

```{r data-collection-year}
dataCollectionYear(metadata)
```

Non-missing `year` values must be numeric or coercible to numeric.

## Instructions

`instructions()` counts contingency-instruction categories once per study. This
prevents multi-condition studies from being counted multiple times for the same
instruction category.

```{r instructions}
instructions(metadata)
```

The mapped metadata must contain `condition_id`, `study_id`, and
`instruction_contingency`.

## Stimulus Modality

`stimModality()` creates a pie chart for unconditioned stimulus modality
(`us_type`) or conditioned stimulus modality (`cs_type`). Slice sizes can count
studies or participants.

```{r stim-modality}
stimModality(
  metadata,
  type = "us_type",
  level = "n_studies"
)

stimModality(
  metadata,
  type = "cs_type",
  level = "n_subjects"
)
```

Missing or blank modality labels are shown as `not reported`. Missing
`n_subjects` values are treated as zero for participant totals.

## Reinforcement Rates

`reinforcementRates()` searches for metadata columns whose names begin with
`reinf`, reshapes those columns to long format, validates numeric values, and
plots whole-number reinforcement-rate counts.

```{r reinforcement-rates}
reinforcementRates(metadata)
```

## SCR Peak Detection Windows

`peakDetectionWindows()` plots SCR baseline, peak-detection, and
trough-detection scoring windows in seconds relative to stimulus onset.

```{r peak-detection-windows}
peakDetectionWindows(
  metadata,
  grouping_variable = "study_id"
)

peakDetectionWindows(
  metadata,
  grouping_variable = "condition_id"
)
```

The function supports current `physio_scr_*` columns and legacy `scr_*` columns.
For trough-to-peak scoring, baseline windows are removed because trough
detection replaces baseline correction.
