---
title: "Participant Summaries"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Participant Summaries}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Load Long-Format Data

```{r setup}
library(fearbase)

data_long <- tibble::tribble(
  ~condition_id, ~study_id, ~paper_cond_id, ~paper_study_id,
  ~participant_id, ~measure, ~value, ~phase, ~stimulus, ~trial,
  "c1", "s1", "pc1", "ps1", "p1", "age", "21", "hab", "cs+", 1,
  "c1", "s1", "pc1", "ps1", "p1", "sex", "m", "hab", "cs+", 1,
  "c1", "s1", "pc1", "ps1", "p2", "age", "24", "hab", "cs+", 1,
  "c1", "s1", "pc1", "ps1", "p2", "sex", "f", "hab", "cs+", 1,
  "c2", "s2", "pc2", "ps2", "p3", "age", "29", "acq", "cs+", 1,
  "c2", "s2", "pc2", "ps2", "p3", "sex", NA_character_, "acq", "cs+", 1,
  "c2", "s2", "pc2", "ps2", "p4", "age", "31", "ext", "cs-", 1
)
```

## Age Plots

`age()` extracts rows where `measure == "age"`, coerces `value` to numeric, and
returns either a stacked histogram or a ridge-density plot.

```{r age-histogram}
age(
  data_long,
  type = "histogram",
  grouping_variable = "study_id"
)
```

```{r age-ridge}
age(
  data_long,
  type = "ridge",
  grouping_variable = "condition_id"
)
```

Supported grouping variables are:

- `condition_id`;
- `study_id`;
- `paper_cond_id`;
- `paper_study_id`.

## Age Descriptives

`age_descriptives()` computes mean, standard deviation, minimum, maximum, and
count for valid age rows. You can compute global or grouped summaries.

```{r age-descriptives}
age_descriptives(data_long)

age_descriptives(
  data_long,
  grouping_variable = c("study_id", "condition_id")
)
```

## Sex Distribution

`sex()` builds a participant-level pie chart from `measure == "sex"` or
`measure == "gender"` rows. Participants without any sex or gender row are
counted as `not reported`, while explicit `NA` sex or gender values are counted
as `NA`.

```{r sex}
sex(data_long)
```

The function accepts common encodings beginning with `m` or `f`, plus common
not-reported labels such as `not reported`, `unknown`, `n/a`, and `nr`.
Contradictory non-missing participant values stop with an error.

## Sample Size by Study

`sample_size_by_study()` counts distinct participants per study-level grouping.
Repeated long-format rows for the same participant do not inflate counts.

```{r sample-size}
sample_size_by_study(data_long)

sample_size_by_study(
  data_long,
  grouping_variable = "condition_id"
)
```

Supported grouping variables are `study_id` and `condition_id`.
