---
title: "Trials, Phases, and Heatmaps"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Trials, Phases, and Heatmaps}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Load Data

```{r setup}
library(fearbase)

data_long <- tibble::tribble(
  ~condition_id, ~study_id, ~paper_cond_id, ~paper_study_id, ~participant_id, ~measure, ~value, ~phase, ~stimulus, ~trial,
  "c1", "s1", "pc1", "ps1", "p1", "scr", "0.3", "hab", "cs+", 1,
  "c1", "s1", "pc1", "ps1", "p1", "scr", "0.4", "acq", "cs+", 1,
  "c1", "s1", "pc1", "ps1", "p1", "us_exp", "70", "acq", "cs-", 1,
  "c1", "s1", "pc1", "ps1", "p2", "scr", "0.2", "ext", "cs+", 1,
  "c1", "s1", "pc1", "ps1", "p2", "us_exp", "30", "ext", "cs-", 1,
  "c2", "s2", "pc2", "ps2", "p3", "scr", "0.5", "acq", "cs+", 1,
  "c2", "s2", "pc2", "ps2", "p3", "scr", "0.6", "acq", "cs+", 2,
  "c2", "s2", "pc2", "ps2", "p4", "us_exp", "20", "ext", "cs-", 1
)

metadata <- tibble::tribble(
  ~condition_id, ~study_id,
  "c1", "s1",
  "c2", "s2"
)

codebook <- tibble::tribble(
  ~attribute, ~abbreviation, ~name,
  "phase", "hab", "habituation",
  "phase", "acq", "acquisition",
  "phase", "ext", "extinction",
  "measure", "scr", "skin conductance response",
  "measure", "us_exp", "US expectancy"
)
```

## Trial Counts per Phase and Participant

`trialsPhaseParticipant()` computes participant-level phase trial counts from
long-format data. It counts the maximum trial index within each participant,
phase, and stimulus stream, then sums across stimulus streams for each
participant-phase pair.

```{r trials-studies}
trialsPhaseParticipant(
  data_long,
  y_axis = "studies",
  grouping_variable = "condition_id",
  cb = codebook
)
```

Use `y_axis = "participants"` to show participant counts at each observed trial
count and fill by the selected grouping variable.

```{r trials-participants}
trialsPhaseParticipant(
  data_long,
  y_axis = "participants",
  grouping_variable = "study_id",
  cb = codebook
)
```

Supported grouping variables are:

- `condition_id`;
- `study_id`;
- `paper_cond_id`;
- `paper_study_id`.

The `int` and `other` phase codes are currently excluded. Phase labels come
from the codebook rows where `attribute == "phase"`.

## Phase Heatmap

`phasesHeatmap()` shows phase co-occurrence across mapped conditions with a
marginal bar chart of phase usage.

```{r phases-heatmap}
phasesHeatmap(data_long, codebook)
```

The function maps identifiers, labels phase abbreviations through the codebook,
removes `int` and `other`, and counts co-occurrence across `condition_id`.

## Measure Heatmap

`measuresHeatmap()` shows outcome-measure co-occurrence across mapped
conditions with a marginal bar chart.

```{r measures-heatmap}
measuresHeatmap(
  data_long,
  metadata,
  codebook
)
```

Measure labels come from the codebook rows where `attribute == "measure"`.

## Reusable Heatmap Helpers

You can use `plot_co_occurrence_heatmap()` and `plot_horizontal_bar()` with
precomputed data.

```{r reusable-helpers}
heatmap_data <- data.frame(
  measure_x = c("SCR", "SCR", "Startle", "Startle"),
  measure_y = c("SCR", "Startle", "SCR", "Startle"),
  value = c(NA_real_, 12, 12, NA_real_)
)

plot_co_occurrence_heatmap(
  heatmap_data,
  x_var = "measure_x",
  y_var = "measure_y",
  value_var = "value",
  diag_na = TRUE
)

bar_data <- data.frame(
  measure = c("SCR", "Startle"),
  used = c(30, 18)
)

plot_horizontal_bar(
  bar_data,
  cat_var = "measure",
  count_var = "used"
)
```
