---
title: "Trials, Phases, and Heatmaps"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Trials, Phases, and Heatmaps}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Load Data

```{r setup}
library(fearbase)
library(readr)

data_long <- read_csv("data/data_long.csv", show_col_types = FALSE)
metadata <- read_csv("data/metadata.csv", show_col_types = FALSE)
codebook <- read_csv("data/codebook.csv", show_col_types = FALSE)
```

## Trial Counts per Phase and Participant

`trialsPhaseParticipant()` computes participant-level phase trial counts from
long-format data. It counts the maximum trial index within each participant,
phase, and stimulus stream, then sums across stimulus streams for each
participant-phase pair.

```{r trials-studies}
trialsPhaseParticipant(
  data_long,
  y_axis = "studies",
  grouping_variable = "condition_id",
  cb = codebook
)
```

Use `y_axis = "participants"` to show participant counts at each observed trial
count and fill by the selected grouping variable.

```{r trials-participants}
trialsPhaseParticipant(
  data_long,
  y_axis = "participants",
  grouping_variable = "study_id",
  cb = codebook
)
```

Supported grouping variables are:

- `condition_id`;
- `study_id`;
- `paper_cond_id`;
- `paper_study_id`.

The `int` and `other` phase codes are currently excluded. Phase labels come
from the codebook rows where `attribute == "phase"`.

## Phase Heatmap

`phasesHeatmap()` shows phase co-occurrence across mapped conditions with a
marginal bar chart of phase usage.

```{r phases-heatmap}
phasesHeatmap(data_long, codebook)
```

The function maps identifiers, labels phase abbreviations through the codebook,
removes `int` and `other`, and counts co-occurrence across `condition_id`.

## Measure Heatmap

`measuresHeatmap()` shows outcome-measure co-occurrence across mapped
conditions with a marginal bar chart.

```{r measures-heatmap}
measuresHeatmap(
  data_long,
  metadata,
  codebook
)
```

Measure labels come from the codebook rows where `attribute == "measure"`.

## Reusable Heatmap Helpers

You can use `plot_co_occurrence_heatmap()` and `plot_horizontal_bar()` with
precomputed data.

```{r reusable-helpers}
heatmap_data <- data.frame(
  measure_x = c("SCR", "SCR", "Startle", "Startle"),
  measure_y = c("SCR", "Startle", "SCR", "Startle"),
  value = c(NA_real_, 12, 12, NA_real_)
)

plot_co_occurrence_heatmap(
  heatmap_data,
  x_var = "measure_x",
  y_var = "measure_y",
  value_var = "value",
  diag_na = TRUE
)

bar_data <- data.frame(
  measure = c("SCR", "Startle"),
  used = c(30, 18)
)

plot_horizontal_bar(
  bar_data,
  cat_var = "measure",
  count_var = "used"
)
```
