---
title: "Fearbase Visualisations"
format: html
toc: true
toc-depth: 4
toc-expand: true
number-sections: true
number-depth: 4
execution:
    cache: true
---

```{r}
#| include: false
#| label: load-packages-data
devtools::load_all() # Load fearbase package
library(tidyverse)

csv_files <- dir(
  "C:\\Users\\calting\\Documents\\GitLab\\FEARBASE-R-Package\\data",
  pattern = "*.csv",
  full.names = TRUE
)

# read all csv files
for (f in csv_files) {
  assign(
    sub(".csv", "", basename(f)),
    read_csv(f),
    envir = .GlobalEnv
  )
}
```

# Participant-related

## Age

### Histogram

I'd remove the histogram from the package due to colours being not differentiable in a meaningful way (see comment in "Introducing the fearbase")

:::: {.column-page layout="[ 50, 50 ]"}

::: {#hist-by-study}

```{r}
#| label: fig-age-hist-study
#| fig-cap: "Age distribution by study (not dataset)."
#| code-fold: true
age(dl = data_long, type = "histogram", grouping_variable = "study_id")
```

:::

::: {#hist-by-condition}

```{r}
#| label: fig-age-hist-condition
#| fig-cap: "Age distribution by condition (= dataset, not study)."
#| code-fold: true
age(dl = data_long, type = "histogram", grouping_variable = "condition_id")
```

:::

::::

### Ridgeline Plots

:::: {.column-page layout-ncol="2"}

::: {#ridge-by-study}

```{r }
#| label: fig-age-ridge-study
#| fig-cap: "Age densities by study (not dataset)."
#| code-fold: true
age(dl = data_long, type = "ridge", grouping_variable = "study_id")
```

:::

::: {#ridge-by-condition}

```{r}
#| label: fig-age-ridge-condition
#| fig-cap: "Age densities by condition (= dataset, not study)."
#| code-fold: true
age(dl = data_long, type = "ridge", grouping_variable = "condition_id")
```

:::

::::


## Sex

::: {#sex-pie-chart}

```{r}
#| label: fig-sex-pie
#| fig-cap: "Number of participants by sex."
#| code-fold: true
sex(dl = data_long)
```

:::

# Study-related

## Year (of Publication or Data Upload)

:::: {.column-page layout-ncol="2"}

::: {#year-of-publication-by-study}
```{r}
#| label: fig-year-publication-by-study
#| fig-cap: "The year a study associated with a Fearbase dataset was published."
#| code-fold: true
data_collection_year(
  md = metadata,
  grouping_variable = "study_id",
  year_of = "publication"
)
```
:::

::: {#year-of-data-upload-by-study}
```{r}
#| label: fig-year-data-upload-by-study
#| fig-cap: "The year a Fearbase dataset was made (publicly) available."
#| code-fold: true
data_collection_year(
  md = metadata,
  grouping_variable = "study_id",
  year_of = "data"
)
```
:::

::::

## Sample Size

:::: {.column-page layout-ncol="2"}

::: {#sample-size-by-study}
```{r}
#| label: fig-sample-size-by-study
#| fig-cap: "Number of participants per study (not dataset)"
#| code-fold: true
sample_size_by_study(dl = data_long, grouping_variable = "study_id")
```
:::

::: {#sample-size-by-condition}
```{r}
#| label: fig-sample-size-by-condition
#| fig-cap: "Number of participants per condition (= dataset)"
#| code-fold: true
sample_size_by_study(dl = data_long, grouping_variable = "condition_id")
```
:::

::::

## Instructions

:::: {.column-page layout-ncol="2" layout-nrow="3"}

::: {#instructions-by-study_with_na}
```{r}
#| label: fig-instructions-study-with-na
#| fig-cap: "Number of studies (not datasets) per instructions (NA included)."
#| code-fold: true
instructions(md = metadata, grouping_variable = "study_id", remove_na = FALSE)
```
:::

::: {#instructions-by-condition_with_na}
```{r}
#| label: fig-instructions-condition-with-na
#| fig-cap: "Number of conditions (= datasets) per instructions (NA included)."
#| code-fold: true
instructions(
  md = metadata,
  grouping_variable = "condition_id",
  remove_na = FALSE
)
```
:::



::: {#instructions-by-study_without_na}
```{r}
#| label: fig-instructions-study-without-na
#| fig-cap: "Number of studies (not datasets) per instructions (NA excluded)."
#| code-fold: true
instructions(md = metadata, grouping_variable = "study_id", remove_na = TRUE)
```
:::

::: {#instructions-by-condition_without_na}
```{r}
#| label: fig-instructions-condition-without-na
#| fig-cap: "Number of conditions (= datasets) per instructions (NA excluded)."
#| code-fold: true
instructions(
  md = metadata,
  grouping_variable = "condition_id",
  remove_na = TRUE
)
```
:::



::: {#instructions-by-study_with_na_by_count}
```{r}
#| label: fig-instructions-study-with_na_by_count
#| fig-cap: "Number of studies (not datasets) per instructions (NA excluded). Sorted by count."
#| code-fold: true
instructions(
  md = metadata,
  grouping_variable = "study_id",
  remove_na = TRUE,
  sort_by_count = TRUE
)
```
:::

::: {#instructions-by-condition_with_na-by-count}
```{r}
#| label: fig-instructions-condition-with-na-by-count
#| fig-cap: "Number of conditions (= datasets) per instructions (NA excluded). Sorted by count."
#| code-fold: true
instructions(
  md = metadata,
  grouping_variable = "condition_id",
  remove_na = TRUE,
  sort_by_count = TRUE
)
```
:::

::::

## Reinforcement Rates

:::: {.column-page layout-ncol="2"}

::: {#reinforcement-rate-by-study}
```{r}
#| label: fig-reinforcement-rate-by-study
#| fig-cap: "Number of Studies (not datasets) per reinforcement rate."
#| code-fold: true
reinforcement_rates(md = metadata, grouping_variable = "study_id")
```
:::

::: {#reinforcement-rate-by-condition}
```{r}
#| label: fig-reinforcement-rate-by-condition
#| fig-cap: "Number of Conditions (= datasets) per reinforcement rate."
#| code-fold: true
reinforcement_rates(md = metadata, grouping_variable = "condition_id")
```
:::

::::

## Stimulus Modalities

::: {.callout-note}
Should we add "n = " to the number of studies/participants in brackets, e.g. auditory (n = 1)?
:::


### US

:::: {.column-page layout-ncol="2"}

::: {#stimulus-modality-us-by-study}
```{r}
#| label: fig-us-modality-by-study
#| fig-cap: "Number of Studies per US modality."
#| code-fold: true
stimulus_modality(md = metadata, type = "us_type", level = "n_studies")
```
:::

::: {#stimulus-modality-us-by-participants}
```{r}
#| label: fig-us-modality-by-subject
#| fig-cap: "Number of Participants per US modality."
#| code-fold: true
stimulus_modality(md = metadata, type = "us_type", level = "n_subjects")
```
:::

::::


### CS

:::: {.column-page layout-ncol="2"}

::: {#stimulus-modality-cs-by-study}
```{r}
#| label: fig-cs-modality-by-study
#| fig-cap: "Number of Studies per CS modality."
#| code-fold: true
stimulus_modality(md = metadata, type = "cs_type", level = "n_studies")
```
:::

::: {#stimulus-modality-cs-by-participants}
```{r}
#| label: fig-cs-modality-by-subject
#| fig-cap: "Number of Participants per CS modality."
#| code-fold: true
stimulus_modality(md = metadata, type = "cs_type", level = "n_subjects")
```
:::

::::


## SCR Scoring-Approaches

::: {.callout-important}
Maybe add "window" plots for other measures than SCR later (e.g. FPS)?
:::

::: {.callout-important}
Rework of scoring approach categories is on our todo.
:::

::: {.callout-note}
*CS Onset* label must be dynamic or relative. Must look further into this.
:::

::: {.callout-note}
An alternative way could be to just show the unique windows, add n = xxx at the left/right of each bar and add a column next to the plot with the study IDs that used the corresponding windows?
:::


:::: {.column-page layout-ncol="2"}

::: {#scr-scoring-approach-by-study}
```{r}
#| label: fig-scr-scoring-approach-by-study
#| fig-cap: "SCR Scoring Approach Parameters by Study."
#| code-fold: true
peak_detection_windows(md = metadata, grouping_variable = "study_id")
```
:::

::: {#stimulus-modality-cs-by-participants}
```{r}
#| label: fig-scr-scoring-approach-by-condition
#| fig-cap: "SCR Scoring Approach Parameters by Condition (= dataset)."
#| code-fold: true
peak_detection_windows(md = metadata, grouping_variable = "condition_id")
```
:::

::::

::: {#stimulus-modality-cs-by-participants}
```{r}
#| label: fig-scr-scoring-approach-dynamic
#| fig-cap: "Unique SCR Scoring Approach Settings. Hover with your mouse over a segment to show Dataset IDs."
#| code-fold: true
peak_detection_windows_dynamic(
  md = metadata,
  grouping_variable = "condition_id"
)
```
:::



## Heatmaps

### Measures

:::{.callout-warning}
This looks messed up only on quarto rendering. When viewing the plot in RStudio or Positron, it looks much better. Will look for a fix, though we might do not want to plot ALL measures at once in a heatmap ;)
A quick and cheap option would be to add an input paramter like "include" that accepts a vector where we can explicitly pass what to plot. In case of many measures this might be cumbersome. The default could be a hardcoded subset, but then again, that would be hardcoded...
:::

:::{.callout-note}
The changes of how ratings are displayed (Valence (US), Expectancy (US)) etc. are **not yet in place**. We might want to think of a way to properly do this manually instead of hardcoding this, e.g. by making some changes to the codebook, maybe?
:::

:::: {.column-screen}
::: {#scr-scoring-approach-by-study}
```{r}
#| label: fig-measures-heatmap
#| fig-cap: "Heatmap for all measures. Note: Measure names are taken from codebook!"
#| code-fold: true
measures_heatmap(
  dl = data_long,
  md = metadata,
  cb = codebook
)
```
:::
::::

### Phases

:::{.callout-note}
Now supports dynamic removal of phases instead of hardcoded exclusion of "int" and "other". Phases to exclude can be supplied as vector, e.g. c("int", "other", "acq").
:::

:::: {.column-screen layout-ncol="2"}

::: {#phases-heatmap-without-int-and-other}
```{r}
#| label: fig-phases-heatmap-without-int-and-other
#| fig-cap: "Heatmap for all phases except 'int' and 'other'. Note: Measure names are taken from codebook!"
#| code-fold: true
phases_heatmap(dl = data_long, cb = codebook, exclude = c("int", "other"))
```
:::

::: {#phases-heatmap-all-phases}
```{r}
#| label: fig-phases-heatmap-all-phases
#| fig-cap: "Heatmap for all phases. Note: Measure names are taken from codebook!"
#| code-fold: true
phases_heatmap(dl = data_long, cb = codebook)
```
:::

::::

### Trials per Phases

Note: The number of trials is calculated as the total number of CSs per phase (e.g. n_trials(acq) = n(CS+) + n(CS-)).

:::: {.column-page layout-ncol="2"}

:::{#studis-per-trial-x-phase}
```{r}
#| label: fig-phases-trials
#| fig-cap: "Number of Studies per trial x phase."
#| code-fold: true
trial_phase_counts(
  dl = data_long,
  y_axis = "studies",
  grouping_variable = "study_id",
  cb = codebook
)
```
:::

:::{#participants-per-trial-x-phase}
```{r}
#| label: fig-phases-trials-participants
#| fig-cap: "Number of Participants per trial x phase."
#| code-fold: true
trial_phase_counts(
  dl = data_long,
  y_axis = "participants",
  grouping_variable = NULL,
  cb = codebook
)
```
:::

::::

# Work in progress & Playground

## Choropleth Map (interactive)

::::{.column-page}

:::{#choropleth-map}

```{r}
#| label: fig-choropleth-map
#| fig-cap: "Number of studies by Country."
#| code-fold: true
#| echo: false
library(sf)
library(ggplot2)
library(ggiraph)
library(patchwork)
library(countrycode)
library(rnaturalearth)
library(rnaturalearthdata)

# Clean fearbase country metadata
metadata <- metadata %>%
  mutate(countryClean = countryname(dataCountry))

# get country counts
fearbase_country_counts <- metadata %>%
  select(countryClean) %>%
  table() %>%
  as_tibble() %>%
  mutate(countryClean = forcats::fct_reorder(countryClean, n))


# Get World Data
world_sf <- ne_countries(scale = "large", returnclass = "sf")
world_sf <- world_sf %>%
  mutate(
    countryClean = countryname(sovereignt, nomatch = NULL)
  )

# add fearbase data to world data
world_sf_complete <- world_sf %>%
  left_join(
    fearbase_country_counts,
    by = c("countryClean" = "countryClean")
  ) %>%
  filter(countryClean != "Antarctica")

# create colour palette
n_countries <- length(unique(fearbase_country_counts$countryClean))
nn = 4
new_palette <- generate_palette(n_countries * nn)
new_colors <- setNames(
  new_palette[seq(0, n_countries * nn, nn)],
  levels(fearbase_country_counts$countryClean)
)

# country bar graph
p1 <- ggplot(
  fearbase_country_counts,
  aes(
    x = reorder(countryClean, n),
    y = n,
    tooltip = n,
    data_id = countryClean,
    fill = countryClean
  )
) +
  geom_col_interactive(data = filter(fearbase_country_counts, !is.na(n))) +
  # geom_text(stat='identity',
  #           aes(label=n),
  #           size = 4,
  #           hjust=-1) +
  guides(fill = "none") +
  scale_fill_manual(values = new_colors) +
  coord_flip() +
  theme(axis.title.x = element_blank(), axis.title.y = element_blank()) +
  scale_y_continuous("Number of datasets", breaks = seq(0, 8))
#p1

p2 <- ggplot() +
  geom_sf(data = world_sf_complete, fill = "lightgrey", color = "lightgrey") +
  geom_sf_interactive(
    data = filter(world_sf_complete, !is.na(n)),
    aes(fill = countryClean, tooltip = countryClean, data_id = countryClean)
  ) +
  coord_sf(crs = st_crs(3857)) +
  theme_void() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    legend.position = "none"
  ) +
  scale_fill_manual(values = new_colors)

combined_plot <- p1 / p2 + plot_layout(heights = c(1, 3))
#combined_plot

# Create the interactive plot
interactive_plot <- girafe(ggobj = combined_plot)
interactive_plot <- girafe_options(
  interactive_plot,
  opts_hover(css = "fill:orange;stroke:black;")
)
interactive_plot
```

:::
::::

## Sex by Study

For now only if both m and f were supplied.

```{r}
#| label: fig-sex-by-study
#| fig-cap: "Number of male and female participants by study. Note: Studies with either sex missing were excluded."
#| code-fold: true
#| warning: false
grouping_variable <- "study_id"

dl <- data_long

data_sex <- data_long |>
  filter(measure == "sex" & !is.na(value)) |>
  mutate(value = .normalize_sex_value(value)) |>
  group_by(.data[[grouping_variable]], "sex") |>
  count(value) |>
  pivot_wider(
    id_cols = any_of(grouping_variable),
    names_from = "value",
    values_from = "n"
  ) |>
  tidyr::drop_na(m, f) |>
  pull(study_id)

data_sex_pyramid <- dl |>
  filter(measure == "sex" & !is.na(value)) |>
  mutate(value = .normalize_sex_value(value)) |>
  group_by(.data[[grouping_variable]], "sex") |>
  count(value) |>
  rename("sex" = "value", "measure" = `"sex"`) |>
  mutate(n_new = dplyr::if_else(sex == "m", -n, n)) |>
  tidyr::drop_na(n_new) |>
  filter(study_id %in% data_sex) |>
  group_by(study_id) |>
  mutate(n_complete = sum(n)) |>
  ungroup()

x_axis_min <- plyr::round_any(min(data_sex_pyramid$n_new), 20, f = floor)
x_axis_max <- plyr::round_any(max(data_sex_pyramid$n_new), 20, f = ceiling)
x_axis_breaks <- seq(x_axis_min, x_axis_max, by = 20)
x_axis_labels <- abs(x_axis_breaks)

data_sex_pyramid |>
  ggplot(aes(x = n_new, y = reorder(study_id, n_complete), fill = sex)) +
  geom_col() +
  labs(
    x = "Number of Participants",
    y = "Study ID"
  ) +
  scale_x_continuous(breaks = x_axis_breaks, labels = x_axis_labels) +
  scale_fill_discrete("Sex", breaks = c("m", "f"), labels = c("male", "female"))
```


## Number of Participants per unique Phase Combination

:::: {.column-page}

::: {#participant-phase-combination}
```{r}
#| label: fig-participants-per-phase-combination
#| fig-cap: "Number of participants per unique phase combination"
#| code-fold: true
`%nin%` <- purrr::negate(`%in%`)


grouping_var <- "condition_id"

study_design <- .apply_mapping_to_study_design(study_design)

study_design <- study_design |>
  filter(name %nin% c("int", "other"))

study_sequences <- study_design %>%
  arrange(grouping_var, orderPosition) %>%
  group_by(.data[[grouping_var]]) %>%
  summarise(
    phase_combination = str_c(name, collapse = " -> "),
    .groups = "drop"
  )

data_long2 <- .apply_mapping_to_long_data(data_long)
data_long2 <- data_long2 |>
  filter(phase %nin% c("int", "other"))


study_counts <- data_long2 %>%
  distinct(.data[[grouping_var]], participant_id) %>%
  count(.data[[grouping_var]], name = "n_participant")

study_summary <- study_sequences %>%
  left_join(study_counts, by = grouping_var)

result <- study_summary %>%
  group_by(phase_combination) %>%
  summarise(
    n_participant = sum(n_participant, na.rm = TRUE),
    n_studies = n(),
    .groups = "drop"
  ) %>%
  arrange(desc(n_participant))

plot_df <- result %>%
  mutate(
    phase_lower = str_to_lower(phase_combination),

    order_group = case_when(
      phase_lower == "hab" ~ 1,
      phase_lower == "hab, acq" ~ 2,
      phase_lower == "hab, acq, ext" ~ 3,
      str_starts(phase_lower, "hab") ~ 4,

      phase_lower == "acq" ~ 5,
      phase_lower == "acq, ext" ~ 6,
      str_starts(phase_lower, "acq") ~ 7,

      phase_lower == "ext" ~ 8,

      TRUE ~ 9
    )
  )

plot_df <- plot_df %>%
  arrange(order_group, phase_combination) %>%
  mutate(
    phase_combination = factor(
      phase_combination,
      levels = phase_combination
    )
  )


p1 <- ggplot(plot_df, aes(x = n_participant, y = phase_combination)) +
  geom_col(fill = "#0032a0") +
  labs(
    x = "Number of Participants",
    y = "Phase Combination"
  ) +
  theme_minimal(base_size = 14) +
  scale_x_continuous(limits = c(0, 300)) +
  #theme(axis.title.y.right = element_text(hjust = -1)) +
  scale_y_discrete(position = "left")


phase_counts <- data_long2 %>%
  filter(phase %nin% c("int", "other")) |>
  tidyr::drop_na(phase) |>
  distinct(study_id, participant_id, phase) %>% # avoid duplicates
  group_by(phase) %>%
  summarise(
    n_participant = n(),
    .groups = "drop"
  ) %>%
  arrange(desc(n_participant))

phase_counts <- phase_counts %>%
  mutate(
    phase_lower = str_to_lower(phase),
    order_group = case_when(
      phase_lower == "hab" ~ 1,
      phase_lower == "acq" ~ 2,
      phase_lower == "ext" ~ 3,
      phase_lower == "rex" ~ 4,
      phase_lower == "rin" ~ 5,
      TRUE ~ 6
    )
  )

phase_counts <- phase_counts %>%
  arrange(order_group, phase_lower) %>%
  mutate(
    phase = factor(phase, levels = phase)
  )

p1
```


```{r}
#| label: fig-participants-per-phase
#| fig-cap: "Number of participants per unique phase."
#| code-fold: true
p2 <- ggplot(phase_counts, aes(x = n_participant, y = phase)) +
  geom_col(fill = "#0032a0") +
  labs(
    x = "Number of Participants.",
    y = "Phase"
  )


p2
```

```{r}
#| label: fig-conditions-per-phase
#| fig-cap: "Number of conditions (= datasets) per unique phase."
#| code-fold: true
p3 <- ggplot(plot_df, aes(x = n_studies, y = phase_combination)) +
  geom_col(fill = "#0032a0") +
  labs(
    x = "Number of Conditions (= datasets)",
    y = "Phase Combination"
  ) +
  theme_minimal(base_size = 14) +
  scale_x_continuous(breaks = seq(0, max(plot_df$n_studies), 2)) +
  #theme(axis.title.y.right = element_text(hjust = -1)) +
  scale_y_discrete(position = "left")

p3
```

:::
::::

## Complex Upset for Participants per Phase (Combination)

```{r}
#| label: fig-participants-per-phase-and-phase-combinations-upset
#| fig-cap: "Phase combinations by condition."
#| code-fold: true
data_long_mapped <- .apply_mapping_to_long_data(data_long)

# Translate compact FEARBASE phase codes to the display names stored in the
# codebook before reshaping. Keeping unmapped codes visible makes exploratory
# local renders robust while still using the codebook as the primary source.
data_long_phases <- data_long_mapped |>
  select("condition_id", "phase") |>
  tidyr::drop_na("condition_id", "phase") |>
  mutate(
    phase = as.character(
      .label_phases_from_codebook(
        phases = .data$phase,
        cb = codebook,
        keep_unmapped = TRUE
      )
    )
  ) |>
  filter(!is.na(.data$phase)) |>
  distinct(.data$condition_id, .data$phase)

phase_order <- c(
  "Habituation",
  "Acquisition",
  "Extinction",
  "Re-Extinction",
  "Reinstatement Test",
  "Reversal",
  "Intervention",
  "Other"
)

observed_phases <- data_long_phases |>
  distinct(.data$phase) |>
  pull(.data$phase)

# Use the requested display order for known phases and append any future
# codebook-derived phase labels after the curated order.
phases <- c(
  phase_order[phase_order %in% observed_phases],
  setdiff(observed_phases, phase_order)
)

# ComplexUpset expects one logical membership column for each set. Here each
# row is one condition and TRUE means that the condition contains that phase.
data_long_upset <- data_long_phases |>
  mutate(in_phase = TRUE) |>
  tidyr::pivot_wider(
    names_from = "phase",
    values_from = "in_phase",
    values_fill = FALSE
  ) |>
  arrange(.data$condition_id)

ComplexUpset::upset(
  data_long_upset,
  phases,
  name = "Phase",
  sort_sets = FALSE,
  base_annotations = list(
    "Number of Participants" = ComplexUpset::intersection_size() +
      ggplot2::ylab("Number of Participants")
  ),
  set_sizes = ComplexUpset::upset_set_size() +
    ggplot2::ylab("Total Number of Participants")
)
```

## Complex Upset for Participants per Phase (Combination), Participant Counts

```{r}
#| label: fig-participants-per-phase-and-phase-combinations-upset-participant-counts
#| fig-cap: "Phase combinations by participant."
#| code-fold: true
# Each row in this version is one participant within one condition. Because
# ComplexUpset counts rows, the intersection size and set size now represent
# participant counts instead of condition counts.
data_long_participant_phases <- data_long_mapped |>
  select("condition_id", "participant_id", "phase") |>
  tidyr::drop_na("condition_id", "participant_id", "phase") |>
  mutate(
    phase = as.character(
      .label_phases_from_codebook(
        phases = .data$phase,
        cb = codebook,
        keep_unmapped = TRUE
      )
    )
  ) |>
  filter(!is.na(.data$phase)) |>
  distinct(.data$condition_id, .data$participant_id, .data$phase)

# Use the same phase vector as the condition-level plot so the two versions
# are directly comparable and keep identical set labels.
data_long_participants <- data_long_participant_phases |>
  mutate(in_phase = TRUE) |>
  tidyr::pivot_wider(
    names_from = "phase",
    values_from = "in_phase",
    values_fill = FALSE
  ) |>
  arrange(.data$condition_id, .data$participant_id)

ComplexUpset::upset(
  data_long_participants,
  phases,
  name = "Phase",
  sort_sets = FALSE,
  base_annotations = list(
    "Number of Participants" = ComplexUpset::intersection_size() +
      ggplot2::ylab("Number of Participants")
  ),
  set_sizes = ComplexUpset::upset_set_size() +
    ggplot2::ylab("Total Number of Participants")
)
```

## Complex Upset for Participants per Measure (Combination), Participant Counts

```{r}
#| label: fig-participants-per-measure-and-measure-combinations-upset-participant-counts
#| fig-cap: "Measure combinations by participant."
#| code-fold: true
metadata_mapped <- .apply_mapping_to_metadata(metadata)

# Metadata marks measure availability with binary family-prefixed columns such
# as physio_scr, rating_fear, and quest_stais. Strip only the family prefix so
# those metadata flags can be matched to the compact codes in data_long$measure.
metadata_measure_lookup <- tibble(
  metadata_measure = names(metadata_mapped)
) |>
  filter(str_detect(.data$metadata_measure, "^(physio|rating|quest)_")) |>
  mutate(
    measure_family = str_extract(.data$metadata_measure, "^[^_]+"),
    measure = str_remove(.data$metadata_measure, "^(physio|rating|quest)_")
  ) |>
  filter(.data$measure %in% unique(data_long_mapped$measure)) |>
  distinct(.data$measure_family, .data$measure, .keep_all = TRUE)

# Pivot the binary metadata flags to condition-measure rows. In the current
# metadata export, used measures are encoded as 1 and unused measures as NA.
metadata_measure_flags <- metadata_mapped |>
  select("condition_id", all_of(metadata_measure_lookup$metadata_measure)) |>
  tidyr::pivot_longer(
    cols = all_of(metadata_measure_lookup$metadata_measure),
    names_to = "metadata_measure",
    values_to = "used_measure"
  ) |>
  filter(.data$used_measure == 1) |>
  left_join(metadata_measure_lookup, by = "metadata_measure") |>
  select("condition_id", "measure", "measure_family") |>
  distinct()

# Use the codebook labels for display. The lookup is intentionally restricted
# to measures identified through the metadata binary flags, excluding general
# participant attributes such as age, sex, gender, and experimental group.
measure_name_mapping <- codebook |>
  filter(.data$attribute == "measure") |>
  select(
    measure = "abbreviation",
    measure_long = "name"
  ) |>
  mutate(
    measure_long = stringr::str_to_title(.data$measure_long)
  )

data_long_measure_data <- data_long_mapped |>
  select("condition_id", "participant_id", "measure") |>
  tidyr::drop_na("condition_id", "participant_id", "measure") |>
  inner_join(metadata_measure_flags, by = c("condition_id", "measure")) |>
  left_join(measure_name_mapping, by = "measure") |>
  filter(!is.na(.data$measure_long)) |>
  distinct(.data$condition_id, .data$participant_id, .data$measure_long)

measure_order <- metadata_measure_lookup |>
  left_join(measure_name_mapping, by = "measure") |>
  mutate(
    measure_family = factor(
      .data$measure_family,
      levels = c("physio", "rating", "quest")
    )
  ) |>
  filter(.data$measure_long %in% data_long_measure_data$measure_long) |>
  arrange(.data$measure_family, .data$metadata_measure) |>
  distinct(.data$measure_long) |>
  pull(.data$measure_long)

# Each row is one participant within one condition. TRUE values indicate that
# the participant has at least one observation for the corresponding measure.
data_long_measures <- data_long_measure_data |>
  mutate(in_measure = TRUE) |>
  tidyr::pivot_wider(
    names_from = "measure_long",
    values_from = "in_measure",
    values_fill = FALSE
  ) |>
  arrange(.data$condition_id, .data$participant_id)

ComplexUpset::upset(
  data_long_measures,
  measure_order,
  name = "Measure",
  sort_sets = FALSE,
  base_annotations = list(
    "Number of Participants" = ComplexUpset::intersection_size() +
      ggplot2::ylab("Number of Participants")
  ),
  set_sizes = ComplexUpset::upset_set_size() +
    ggplot2::ylab("Total Number of Participants")
)
```
